---
title: "Group 35 CA pollution burden"
author: "Karina Johnston, Caroline Shepherd"
date: "11/15/2021"
output: html_document
---

```{r setup, include = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Attach packages
library(tidyverse)
library(janitor)
library(here)
library(hexbin)  # new package to use `geom_hex()` function
```

This file is an exploration of California environmental justice data using CalEnviroScreen 3.0 scores (CES 3.0), or the California Communities Environmental Health Screening Tool (https://caleja.org/wp-content/uploads/2015/09/CES3.0-5.pdf). CES 3.0 uses 20 indicators to provide a statewide ranking of all 8,000 census tracts in California. Communities with high total CES 3.0 scores have high cumulative pollution burdens and highly vulnerable population characteristics that cause pollutants to become more deadly.

```{r}
# Read in the CA Environmental Justice csv and clean column names

ces <- read_csv(here("data", "ces3demographics.csv")) %>% 
  clean_names()
```


```{r, fig.cap = "**Figure 1.** Relationship between CalEnviroScreen pollution burden scores (3.0) and white percent of the community where high CES scores indicate high pollution burden."}

# Create hexbin graph to show the relationship between white % and CalEnviroScreen pollution burden score

ggplot(data = ces, aes(x = white_percent, y = ces_3_0_score)) +
  geom_hex() +
  labs(x = "White Population (%)", 
       y = "CalEnviroScreen 3.0 Score",
       title = "Pollution Burden Scores and White Population (Johnston & Shepherd)") +
  scale_fill_gradient(low = "yellow", high = "red") + # Updates the color fill pattern
  theme_minimal() # updates the theme to show cleaner output graph

```

Figure 1 shows that communities with a high percentage of white populations tend to have lower pollution burdens overall, and areas with low percentages of white populations tend to have higher pollution burdens overall. There appears to be a negative correlation between percentage of white populations and pollution burden, which would be a good next step to test statistically.

Data Citation: Rodriquez, Matthew and Ziese, Lauren. 2017. CalEnviroScreen 3.0. California Environmental Protection Agency. https://oehha.ca.gov/calenviroscreen/report/calenviroscreen-30

